---
title: "Toturial"
author: "chenxi"
date: "2024-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Add data

```{r}
gene_names <- paste0("Gene", 1:10)
cell_names <- paste0("Cell", 1:10)
expression_matrix <- matrix(rpois(10 * 10, lambda = 5), nrow = 10, ncol = 10)
rownames(expression_matrix) <- gene_names
colnames(expression_matrix) <- cell_names
sparse_matrix <- as(expression_matrix, "sparseMatrix")
seurat_object <- CreateSeuratObject(counts = sparse_matrix, project = "ExampleProject")
seurat_object@meta.data <- Tcell.meta
```

##COIVD

```{r}


covid.data <- readRDS("COVID2024-01-17105618.rds")
covid.data$dataset
covid.data$sampleID
covid.data$cellratio <- covid.data$num/covid.data$allnum
covid.data$tissue
covid.data$severity
covid.data <- covid.data[,c("sampleID","celltype","cellratio","dataset","tissue","severity")]
usethis::use_data(covid.data,overwrite = T)


```



## MNP

```{r}


MNP.seurat_object <- readRDS("I:\\BaiduSyncdisk\\DT20230523已备份\\DT20230801\\20231104CWAS\\20240605DrugALLcal\\MNP2021_MNP_Verse.RDS")

MNP.metadata <- MNP.seurat_object@meta.data
 MNP.seurat_object <- seurat_object
MNP.seurat_object@meta.data <- MNP.metadata
  
usethis::use_data(MNP.seurat_object,MNP.metadata,overwrite = T)

sparkle.data <- cwas_build_model_data(inputdata = MNP.metadata,Sample = "orig.ident",Phenotype = "Status",Celltype ="Clusters",Group ="Tissue",Subgroup = "Study.No",Control_label = "Healthy",Disease_label = "Cancer")
 

```


##Tcell

```{r}

Tcell.metadata <- readRDS("data/metaInfo/panC.freq.all.meta.tb.rds")
Tcell.metadata <- as.data.frame(Tcell.meta)
Tcell.seurat_object <- seurat_object
Tcell.seurat_object@meta.data <- Tcell.metadata
  
usethis::use_data(Tcell.seurat_object,Tcell.metadata,overwrite = T)

sparkle.data <- cwas_build_model_data(inputdata = Tcell.seurat_object,Sample = "patient",Phenotype ="treatment",Celltype = "meta.cluster.coarse",Group = "cancerType",Subgroup = "dataset.old",Control_label = "baseline",Disease_label = "post.treatment")
cwas_pcadraw(MoMac.cwas.data)
 
```


# How to use SPARKLE

## Install SPARKLE

```{r}
devtools::install_github("chenxi199506/SPARKLE")
```

## Step 1: Input data

SPARKLE supports multiple input formats. Users can directly input single-cell Seurat objects,Seurat metadata, or dataframes with calculated cell proportions.


### Input formats 1: Seurat objects

```{r}
seurat_object <-SPARKLE::seurat_object 
 


```


### Input formats 2: Seurat metadata

```{r}

metadata <-SPARKLE::metadata 
metadata <- metadata[1:7000,]  
seurat_object <- CreateSeuratObject(counts = expression_matrix <- matrix(rpois(1000 * 100, lambda = 5), nrow = 1000, ncol = 100), project = "ExampleProject")
seurat_object@meta.data <- metadata
sparkle.data <- cwas_build_model_data(inputdata = seurat_object,Sample = "Patient.No",Phenotype = "Status",Celltype ="Clusters",Group ="Tissue",Subgroup = "Study.No",Control_label = "Healthy",Disease_label = "Cancer")
 

```

### Input formats 3: data frame with cell rate information

```{r}

library(SPARKLE)

covid.data.ratio <-SPARKLE::covid.data 

sparkle.data <- cwas_build_model_data(inputdata = covid.data.ratio,Sample = "sampleID",Phenotype = "severity",Celltype ="celltype",Group ="tissue",Subgroup = "dataset",Control_label = "control",Disease_label = "severe/critical",Cellrate ="cellratio" )

```

## Step 2:Data exporlation (Optional)

```{r}

cwas_pcadraw(sparkle.data) # PCA plot for data exploration
 
cwas_classic_comparision(sparkle.data) #wilcoxon rank sum test and plot

```


## Step 3: Calculation for all model
 

```{r}
cwas.test.data

all.models <- cwas_allmodel_cal2(cwas.test.data)### calculation for all model
#all.models <- cwas_allmodel_cal(sparkle.data)### calculation for all model

cwas_allmodel_heatmap(all.models)# Plot all model AIC and Pvalue

best.model <- cwas_autoselected_model(all.models)# Select the best model and plot
 

```

